Citation Analysis of Computer Science Publications

### 2019-12-06 Scientometrics/comp.tex
cited the Borner paper for why we use conductance, and otherwise just minor edits in the first 6 pages

\paragraph{Clustering by Direct Citation} Graclus is a spectral graph clustering package that computes normalized cut and ratio association for a given undirected graph~\cite{graclus_2007}.  We used v1.2 in our experiments. The \emph{comp} dataset was formatted as an undirected graph, stored in a file with a header line indicating the number of nodes and edges, and used as input to Graclus, which requires the number of clusters to be formed as an input parameter. In preliminary experiments, we varied the number of clusters to be formed between 10 and 50 clusters (data not shown).  At 20 clusters, the most evenly sized set of clusters resulted with the largest cluster containing roughly 10 times the number of nodes in the smallest one~\cite{traag_louvain_2019}.  Consequently, we generated 20 clusters, labeled 0-19, using Graclus (Fig~\ref{fig:heatmap}). The quality of clustering was evaluated by measuring conductance as defined in Shun et al.~\cite{shun_parallel_2016}. 
%George, Tandy - add text about why conductance is good, and that small values are better
Data in \emph{comp} were clustered into 18, 20, or 22 clusters and in each case the highest numbered cluster (17, 19, or 21) had the greatest conductance value and the smallest number of nodes, suggesting that it effectively served as a container for `left over publications' during the Graclus process. Thus, we did not consider cluster 19 (the 20th cluster) when interpreting our results.  The remaining 19 clusters had conductance values ranging from 0.09 to 0.25 with a median conductance of 0.15.

@@

We applied variable clustering with initial parameters of threshold ($t$) = median normalized co-citation frequency (quantile=0.5), increment $i = 0.1$, and maximum cluster size, $mcs=200$. Thus, at start, all edges below the median normalized co-citation frequency were deleted. Clusters were formed by assembling connected components from each co-cited pair beginning with the heaviest edge weight. Clusters below size 100 were retained and any cluster larger than 200 nodes was carried over to the next round. The threshold, $t$, was then incremented by 0.1 and the process repeated while progressively incrementing $t$.  We used a bi-phasic approach where in which $t$ ranged from 0.5-0.9, after which $i$ was reduced to 0.01 for the range $0.9 \leq t \leq 0.99$. A final threshold of $t$=0.999 was applied to break the single remaining large cluster. All clusters generated thus, contained  less than 100 nodes.  Using this approach, 22,232 clusters containing 84,591 nodes were generated. Clusters containing only 2 nodes were then discarded bringing the total number of clusters down to 10,298.

>>>>

\paragraph{Clustering by Direct Citation} Graclus is a spectral graph clustering package that computes normalized cut and ratio association for a given undirected graph~\cite{graclus_2007}.  We used v1.2 in our experiments. The \emph{comp} dataset was formatted as an undirected graph, stored in a file with a header line indicating the number of nodes and edges, and used as input to Graclus, which requires the number of clusters to be formed as an input parameter. In preliminary experiments, we varied the number of clusters to be formed between 10 and 50 clusters (data not shown).  At 20 clusters, the most evenly sized set of clusters resulted with the largest cluster containing roughly 10 times the number of nodes in the smallest one~\cite{traag_louvain_2019}.  Consequently, we generated 20 clusters, labeled 0-19, using Graclus (Fig~\ref{fig:heatmap}).
We used conductance, as defined in Shun et al.~\cite{shun_parallel_2016}, to  evaluate clustering (smaller is better), noting that
conductance has been found to be ``the stand-alone quality metric that best indicates performance on the information recovery metrics" \cite{emmons2016analysis}.
% The quality of clustering was evaluated by measuring conductance as defined in Shun et al.~\cite{shun_parallel_2016}. 
%George, Tandy - add text about why conductance is good, and that small values are better
The data in \emph{comp} were clustered into 18, 20, or 22 clusters and in each case the highest numbered cluster (17, 19, or 21) had the greatest conductance value and the smallest number of nodes, suggesting that it effectively served as a container for `left over publications' during the Graclus process. Thus, we did not consider cluster 19 (the 20th cluster) when interpreting our results.  The remaining 19 clusters had conductance values ranging from 0.09 to 0.25 with a median conductance of 0.15.

@@

We applied variable clustering with initial parameters of threshold ($t$) = median normalized co-citation frequency (quantile=0.5), increment $i = 0.1$, and maximum cluster size, $mcs=200$. Thus, at start, all edges below the median normalized co-citation frequency were deleted. Clusters were formed by assembling connected components from each co-cited pair beginning with the heaviest edge weight. Clusters below size 100 were retained and any cluster larger than 200 nodes was carried over to the next round. The threshold, $t$, was then incremented by 0.1 and the process repeated while progressively incrementing $t$.  We used a bi-phasic approach where in which $t$ ranged from 0.5-0.9, after which $i$ was reduced to 0.01 for the range $0.9 \leq t \leq 0.99$. A final threshold of $t$=0.999 was applied to break the single remaining large cluster. % All generated clusters  using this process contained  less than 100 nodes.  
Using this approach, 22,232 clusters containing 84,591 nodes were generated, each containing less than 100 nodes. Clusters containing only 2 nodes were then discarded bringing the total number of clusters down to 10,298.
